---
title: "correlation"
author: "Miley Osiris God of Death"
date: "10/17/2020"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Portfolio correlation

When securities in a portfolio show high correlation, risk is amplified since a market shock for any reason is likely to have a large effect on all the securities. Add tickers below to see portfolio correlation.
```{r}
my.tickers <- c("DX-Y.NYB", "BTC-USD", "GC=F")

```

## Back-end

These packages help with data science, visualization, and getting financial data from Yahoo Finance's API.  They'll only install once, once they are in your R environment you just have to load them in future sessions.
```{r}
 # Install and load required packages
list.of.packages <- c("tidyverse", "caret", "BatchGetSymbols", "GGally")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, dependencies = TRUE, repos='http://cran.us.r-project.org')

lapply(list.of.packages, require, character.only = TRUE)
```
## Get stock data. 
You can look at different timescales by changing the first and last date and the frequency.
```{r}
first.date <- Sys.Date() - 1825
last.date <- Sys.Date()
freq.data <- 'daily'


stock.data <- BatchGetSymbols(tickers = my.tickers %>% append("^GSPC"), 
                         first.date = first.date,
                         last.date = last.date, 
                         freq.data = freq.data,
                         cache.folder = file.path(tempdir(), 
                                                  'BGS_Cache') ) 

condensed.data <- stock.data$df.tickers %>% 
  select(ticker, ref.date, price.close) %>% 
  spread(ticker, price.close) %>% 
  rename("S&P 500" = `^GSPC`)

```
## Visualizations

1. Correlation matrix - shows how much each security correlates with each other security. This is a cool chart but don't bother if you have more than 5 or so securities - it's gonna be unreadable

```{r}
condensed.data %>% select(-ref.date) %>% ggpairs()
```

2. Stacked line chart. Securities are centered and scaled then plotted-- you can see how they vary over time. If your securities are all moving in the same direction at the same time, it shows high correlation and you may want to diversify your portfolio
```{r}

# Function to scale each column indvidually
scale.data.frame <- function(dfr) {
    if (!is.data.frame(dfr)) {stop("dfr must be a data frame")}
    x <- dfr
    cols <- sapply(dfr, is.numeric)
    scaledvars <- scale.default(dfr[, cols]) 
    x[, cols] <- scaledvars
    return(x)
}

# run scaling function, gather for plotting
scaled <- condensed.data %>% 
  scale.data.frame() %>% 
  gather(key = "ticker", value = "scaled.price", -ref.date)

# Plot
scaled %>% ggplot(aes(x = ref.date, y = scaled.price, color = ticker)) +
  geom_line(size = 2, alpha = .5) +
  labs(title = "Centered and scaled price movements", x = "Date", y = "Scaled price")

```
